\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2016-01-24 overview.tex
updates

In this work, we are proposing a solution between the above two extremes: test the implementation of one or more components against a modeled \psharp environment. The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, and achieve this in a much reduced state-space. Testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp. However, developers already spend significant time in building test suites for distributed systems prior to deployment. The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}). In principle, our methodology is \emph{not specific} to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.

@@

\psharp machines run concurrently with each other, each executing an event handling loop that dequeues the next event from the queue and handles it by invoking the registered action. An action might transition the machine to a new state, create a new machine, send an event to another machine, access a field or call a method. In \psharp, a send operation is \emph{non-blocking}; the event is simply enqueued into the queue of the target machine, which will dequeue and handle the event concurrently. All this functionality is provided in a lightweight runtime library, built on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.

@@

We do not model the server component since we want to test its actual implementation. The server is wrapped inside a \psharp machine, which is responsible for: (i) sending events via the \texttt{PSharp.Send(...)} method, instead of the real network; and (ii) delivering received events to the wrapped component. We model the SNs so that they store data in memory rather than on disk (which could be inefficient during testing). We also model the client so that it can drive the system by repeatedly sending a \texttt{ClientReq} and then waiting for an \texttt{Ack}. Finally, we model the timer so that \psharp takes control of all time-related nondeterminism in the system. This allows the \psharp runtime to control when a \texttt{Timeout} event will be sent to the SNs during testing, and systematically explore different schedules.

@@

In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example is correct. The details of how \psharp was used to model and test real distributed storage systems in Microsoft are covered in \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric}. Interested readers can also refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.\ Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}).

>>>>

In this work, we are proposing a solution between the above two extremes: test the real implementation of one or more components against a modeled in \psharp environment. The benefit of our approach is that it can detect bugs in the \emph{actual} implementation by exploring a much reduced state-space. Note that testing with \psharp does not come for free; developers have to invest effort and time into building a test harness using \psharp. However, developers already spend significant time in building test suites for distributed systems prior to deployment. The \psharp approach augments this effort; by investing time in modeling the environment, it offers dividends by finding more bugs (see \S\ref{sec:eval}). In principle, our methodology is \emph{not} specific to \psharp and the .NET framework, and can be used in combination with any other programming framework that has equivalent capabilities.

@@

\psharp machines run concurrently with each other, each executing an event handling loop that dequeues the next event from the queue and handles it by invoking the registered action. An action might transition the machine to a new state, create a new machine, send an event to a machine, access a field or call a method. In \psharp, a send operation is \emph{non-blocking}; the event is simply enqueued into the queue of the target machine, which will dequeue and handle the event concurrently. All this functionality is provided in a lightweight runtime library, built on top of Microsoft's Task Parallel Library~\cite{leijen2009tpl}.

@@

We do not model the server component since we want to test its actual implementation. The server is wrapped inside a \psharp machine, which is responsible for: (i) sending events via the \texttt{PSharp.Send(...)} method, instead of the real network; and (ii) delivering received events to the wrapped component. We model the SNs so that they store data in memory rather than on disk (which could be inefficient during testing). We also model the client so that it can drive the system by repeatedly sending a nondeterministically generated \texttt{ClientReq}, and then waiting for an \texttt{Ack} message. Finally, we model the timer so that \psharp takes control of all time-related nondeterminism in the system. This allows the \psharp runtime to control when a \texttt{Timeout} event will be sent to the SNs during testing, and systematically explore different schedules.

@@

In \S\ref{sec:overview:safety} and \S\ref{sec:overview:liveness}, we discuss how safety and liveness specifications can be expressed in \psharp to check if the example system is correct. The details of how \psharp was used to model and test real distributed storage systems in Microsoft are covered in \S\ref{sec:vnext}, \S\ref{sec:migrating} and \S\ref{sec:fabric}. Interested readers can also refer to the \psharp GitHub repository\footnote{\url{https://github.com/p-org/PSharp}} to find a manual and samples (e.g.\ Paxos~\cite{lamport1998part} and Raft~\cite{ongaro2014raft}).

### 2016-01-24 vnext.tex
updates

\textbf{Real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maps extents to their hosting ENs. It is updated upon receiving a periodic sync report from an EN. Recall that a sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update the ExtMgr's possible out-of-date view of the EN with the ground truth. The \texttt{ExtentNodeMap} maps ENs to their latest heartbeat times.

@@

Intercepting all network messages and dispatching them through \psharp is important for two reasons. First, it allows \psharp to systematically explore the interleavings between asynchronous event handlers in the system. Second, the modeled network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable (not shown in this example).

>>>>

\textbf{Real Extent Manager}.
The real ExtMgr (see Figure~\ref{fig:extentmanager}) contains two data structures: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maps extents to their hosting ENs. It is updated upon receiving a periodic sync report from an EN. Recall that a sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update the ExtMgr's possibly out-of-date view of the EN with the ground truth. The \texttt{ExtentNodeMap} maps ENs to their latest heartbeat times.

@@

Intercepting all network messages and dispatching them through \psharp is important for two reasons. First, it allows \psharp to systematically explore the interleavings between asynchronous event handlers in the system. Second, the modeled network engine could leverage the support for controlled nondeterministic choices in \psharp, and choose to drop the messages in a nondeterministic fashion, in case emulating message loss is desirable (not shown in this example).
